
# Reproducible Research: Peer Assessment 1
## Basic Global Settings
```{r settings}
echo = TRUE          # Always make code visible
options(scipen = 1)  # Turn off scientific notations for numbers
```

## Plotting System
```{r plotting systems, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
```

## Loading and preprocessing the data
### 1. Load the data (i.e. read.csv())
```{r load data, results='markup', warning=TRUE, message=TRUE}
if(!file.exists("activity.csv")){
    unzip("activity.zip")
}
activity <- read.csv("activity.csv")
```
### 2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r process}
# No processing needed
```

## What is mean total number of steps taken per day?
### 1. Calculate the total number of steps taken per day
#### ignore the missing values in the dataset
```{r calculate}
stepsDay <- aggregate(activity$steps, list(nbr=as.numeric(activity$date), date=activity$date), FUN=sum, na.rm=TRUE)
```

### 2. Make a histogram of the total number of steps taken each day
```{r histogram}
ggplot (stepsPerDay, aes(nbr, x)) + 
  geom_bar(stat="identity", width=0.5) +
  labs(title = "Histogram of Total Number of Steps Taken Each Day", x = "Day", y = "Total Number of Steps Taken Each Day")
```

### 3. Calculate and report the mean and median of the total number of steps taken per day
```{r report mean and median}
#stepsMean <- mean(stepsPerDay)
#stepsMedian <- median(stepsPerDay)
```
* Mean: `r stepsMean`
* Median:  `r stepsMedian`


```
