
# Reproducible Research: Peer Assessment 1
## Basic Global Settings
```{r settings}
echo = TRUE          # Always make code visible
options(scipen = 1)  # Turn off scientific notations for numbers
```

## Plotting System
```{r plotting systems, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
```

## Loading and preprocessing the data
### 1. Load the data (i.e. read.csv())
```{r load data, results='markup', warning=TRUE, message=TRUE}
if(!file.exists("activity.csv")){
    unzip("activity.zip")
}
activity <- read.csv("activity.csv")
```
### 2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r process}
# No processing needed
```

## What is mean total number of steps taken per day?
### 1. Calculate the total number of steps taken per day
#### ignore the missing values in the dataset
```{r calculate}
stepsDay <- aggregate(activity$steps, list(nbr=as.numeric(activity$date), date=activity$date), FUN=sum, na.rm=TRUE)
```

### 2. Make a histogram of the total number of steps taken each day
```{r histogram}
ggplot (stepsDay, aes(nbr, x)) + 
  geom_bar(stat="identity", width=0.5) +
  labs(title = "Histogram of Total Number of Steps Taken Each Day", x = "Day", y = "Total Number of Steps Taken Each Day")
```

### 3. Calculate and report the mean and median of the total number of steps taken per day
```{r report mean and median}
stepsMean <- mean(stepsDay$x)
stepsMedian <- median(stepsDay$x)
```
* Mean: `r stepsMean`
* Median:  `r stepsMedian`

## What is the average daily activity pattern?
### 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r average}
avg <- aggregate(x=list(steps=activity$steps), by=list(interval=activity$interval),  FUN=mean, na.rm=TRUE)
ggplot(avg, aes(interval, steps)) + 
  geom_line(size = 0.5) + 
  labs(title = "Time Series Plot of the 5-minute Interval", x = "5-minute Interval", y = "Average Number of Steps Taken")
```

### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r max steps}
maxSteps = avg[avg$steps == max(avg$steps), ]
```

* 5-Minute Interval: `r maxSteps$interval` Has Maxium Steps of: `r maxSteps$steps`




